### 2023-03-24 discussion.qmd
update

The development of the UVP6 classification model started without a data set of labeled UVP6 images. Therefore, a significant part of my work was to test and prepare a fully functional data preparation and model training pipeline for the upcoming UVP6 data set using UVP5 images. In the following sections, I will describe a brief historic of what has been done prior to the acquisition and human labeling of UVP6 images and discuss about the use and perspectives on the embedded classification.

@@

```{r, kerguelen, fig.width=9, fig.height=8}
#| label: fig-kerguelen
#| fig-cap: $\textit{Calanoida}$, $\textit{Chaetognatha}$, detritus, puff and tuff (i.e. $\textit{Trichodesmium}$) concentrations in February-March 2023 on the Kerguelen Plateau.

@@

Unless we find a cheap and efficient (low power) way to send UVP6 images via Iridium communication, we are stuck with the issue of embedded classification. In @sec-uvp6, I showed that CNNs usually outperform features-based classifiers. However, the field of computer vision is constantly evolving with the arrival of new technologies that could help improve the automatic classification of zooplankton images. Transformers [@Vaswani2017-il], also referred to as Foundation Models in @Bommasani2021-kb are the latest cutting-edge tool in the field of computer vision and natural language processing [e.g. @Liu2021-jq; @Li2022-qt]. The structure of transformers relies on attention mechanisms able to draw global dependencies between input and output [@Vaswani2017-il]. In other words, it is able to learn dependencies between distant positions in a network architecture in constrast to CNNs. As an example, @Li2022-qt used a vision transformer model combining a CNN and a transformer to identify apple diseases. In their model, the CNN is used to extract the global features of the input image while the transformer extracts local features of the disease region to help the CNN see better. How does it help the CNN? In the case of apple images with different backgrounds (different weather conditions, brightness changes, other objects in the field of view), the CNN could be "distracted" hence the transformer is able to focus the attention of the CNN on the region of interest (the disease region) and not on irrelevant features such as the weather. The use of transformers for automatic zooplankton classification has just begun [@Kyathanahally2022-ma] but it should certainly be explored further in the near future.

@@

### The straightforward approach

@@

```{r, uvp traps th234 map}
#| label: fig-uvp-trap-mapping
#| fig-cap: Spatial distribution of sediment traps (A), $^{234}Th$ (B), UVP5 (C) and their spatio-temporal match-ups (D)

@@

The idea was based on the assumption that carbon fluxes could be similar in distinct regions of the ocean if those regions had similar environmental conditions thus removing spatial (horizontal) and temporal constraints on match-ups. For this, we tried different combinations of environmental variables such as surface temperature, salinity, nutrients, combination of nutrients, Chla, $b_{bp}$ or water column properties such as integrated Chla content, oxygen minimum and MLD from monthly climatologies. Then, we used those environmental variables to build a 3D environmental space using a 3D principal component analysis (PCA) decomposition and we projected UVP5 data and trap/$^{234}Th$ in that space based on the environmental characteristics from the respective coordinates and month of each UVP5 and trap/$^{234}Th$ profiles.

@@

This method increased the number of match-ups to \~7800 however the minimization did not find a clear local minimum in the solution space (@fig-solution-space). Instead of having a clear local minimum, @fig-solution-space shows what looks like a 2D solution plane. In addition, I think the choice of $A$ and $D$ (or $B$, $B = 2D - 1$ ) is not constrained enough. In Figure 5 of @Guidi2008-vp, I think it is difficult to really say that there is a local minimum. The light gray color area (low residual error) encompasses a wide region of $A$ and $B$ values where their jackknife procedure found local minima with ($A$, $B$) pairs such as (50, 2) or (150, 5). While $A$ is seen as a constant with no physical meaning, the value of $B$ should really be questioned since it is related to the fractal dimension $D$ of particles and aggregates. Fractal dimensions should not exceed 3 (physical limit) but they have been estimated down to 1.1-1.3 [@Alldredge1988-lx]. In addition, it is possible to obtain coherent POC fluxes by tuning the $A$ value for a fixed $B$ therefore I have my doubts about this method because I am not sure $A$ and $B$ are trully independent.

@@

So far, we have seen that the match-up quest was not enough in the geographical space and not working in the environmental space but what about a method that does not need any match-ups? With XGBoost, it is possible to create a regression model to predict carbon fluxes using environmental variables and geographical variables (depth) as predictors (i.e. explanatory variables) and trap/$^{234}Th$ POC fluxes as target variables. Once the model is trained, it can be applied to the entire data set of PSDs (\> 1.5 million) to derive POC fluxes. Our model had a $R^{2}$ of 0.47 with a percentage of explained variance dominated by depth (54%), surface Chla (16%), surface temperature (13%) and the oxygen minimum (10%). The minimization of @Guidi2008-vp applied on all true PSD and predicted POC fluxes led to the same kind of solution space as shown in @fig-solution-space.

Instead of directly predicting carbon fluxes using environmental variable, we could focus on predicting PSD slope or PSD. In a recent paper, @Clements2022-oq used physical and biological variables (see their Table 1) to predict global maps of PSD slope and biovolume at the base of the EZ using UVP5 data and a random forest algorithm. @Clements2022-oq highlighted the importance of Chla (i.e. proxy of biomass), oxygen (i.e. proxy of respiration) and MLD as predictors in their random forest model. They also emphasized the importance of including phytoplankton functional groups and abundance as key controlling factors on both the PSD slope and the biovolume in future models due to their impact on the downward POC flux (i.e. aggregation/disaggregation processes). The study of @Denvil-Sommer2022-oi used both random forest and XGBoost models to predict small and large POC concentrations using geographical variables (depth, latitude and longitude), physical variables (incident light, MLD and temperature), chemical (phosphate and nitrate) and biological variables (Chla and 12 Plankton Functional Types (PFTs) such as diatoms or microzooplankton). @Denvil-Sommer2022-oi show that for small POC, the most important predictors are the depth level, temperature, microzooplankton and phosphate while the large POC is mostly driven by depth, temperature, MLD, microzooplankton, phaeocystis (another PFT), phosphate and Chla averaged over the MLD. Overall, the results of these studies suggest that it is possible to link the surface environmental and ecosystem structure to the POC distribution in the ocean interior.

@@

I used three methods to derive POC fluxes from UVP5 PSDs. The main issues I encountered were related to the number and accuracy of match-ups as well as the choice of environmental variables to link the surface ocean to the POC distribution below. With the deployment of new BGC-Argo floats equipped with both the UVP6 and the OST as well as the core biogeochemical variables (e.g. Chla, $b_{bp}$, incident light) of the BGC-Argo program, we will hopefully be able to better constraint the PSD-POC relationship however we will probably need a few more years to acquire enough profiles in different oceanic regions to better understand the key drivers of the downward POC flux.

## Continuous vertical sequestration

Using the CONVERSE approach, we have estimated that the global carbon sequestration (\> 100 years) flux by the BCP is 0.9-2.6 Pg C y$^{-1}$. This is 2-3 and 3-6 times larger than POC fluxes at, respectively, 1000 and 2000 m. CONVERSE sequestration flux estimates are also up to 4 times larger than those from studies assuming that sequestration occurs below a fixed reference depth [@Honjo2008-kn, @Guidi2015-sj; @Henson2012-ck]. Here below, I highlight some key points and ways to improve our estimations.

@@

Our study on the sequestration flux by the BCP also shows that the contribution of the vertical migration and mixing pumps contribute at least to 21-34% of the sequestration flux by the BCP. However, those two pumps could probably be better estimated with a multi-instruments approach [@Claustre2021-tx].

@@

With CONVERSE, we estimate that the mixing pump contributes to 11-23% of the sequestration flux by the BCP. However, we were not able to distinct the contribution of the three identified pathways composing the mixing pump due to lacking information in the literature. This could be improved by using the observational framework developed in @Claustre2021-tx by combining satellite, ship-based, mooring and AUV data. BGC-Argo floats have shown that they are adapted to study the detrainment of mixed-layer particles [@DallOlmo2014-ot, @DallOlmo2016-ex] and so have gliders for the eddy subduction pump [@Stukel2017-lc; @Omand2015-mf]. Nonetheless, the study of the large-scale subduction pump will be more challenging because it will require a larger number of measurements made at the scale of ocean basins and over annuals cycles [@Claustre2021-tx], hence the need for a multi-instruments approach to better understand it.

@@

With CONVERSE, we estimate that the vertical migration pump contributes to 9-12% (diel) and 0.1-0.4% (seasonal). Our estimation of the seasonal migration pump is limited to the northern North Atlantic hence it is underestimated. The

dire qu'il faut améliorer la physical et la VMP (diel et saisonnière) -\> mettre le paquet sur ARGO (et les gliders pour la physical) + papier d'Hervé de 2021. + préciser qye la migration saisonnire est sous-évaluée. + acoustique? passsive acoustique? parler de nowicky pour la migration saisonnière en disant qu'il y a aussi les poissons qui peuvent migrer etc

>>>>

The development of the UVP6 classification model started without a data set of labeled UVP6 images. Therefore, a significant part of my work was to test and prepare a fully functional data preparation and model training pipeline for the upcoming UVP6 data set using UVP5 images. In the following sections, I will describe a brief history of what has been done prior to the acquisition and human labeling of UVP6 images and discuss about the use and perspectives on the embedded classification.

@@

```{r, kerguelen, fig.width=8, fig.height=10}
#| label: fig-kerguelen
#| fig-cap: $\textit{Calanoida}$, $\textit{Chaetognatha}$, detritus, puff and tuff (i.e. $\textit{Trichodesmium}$) concentrations in February-March 2023 on the Kerguelen Plateau.

@@

Unless we find a cheap and efficient (low power) way to send UVP6 images via Iridium communication, we are stuck with the issue of embedded classification. In @sec-uvp6, I showed that CNNs usually outperform features-based classifiers. However, the field of computer vision is constantly evolving with the arrival of new technologies that could help improve the automatic classification of zooplankton images. Transformers [@Vaswani2017-il], also referred to as Foundation Models in @Bommasani2021-kb are the latest cutting-edge tool in the field of computer vision and natural language processing [e.g. @Liu2021-jq; @Li2022-qt]. The structure of transformers relies on attention mechanisms able to draw global dependencies between input and output [@Vaswani2017-il]. In other words, it is able to learn dependencies between distant positions in a network architecture in constrast to CNNs. As an example, @Li2022-qt used a vision transformer model combining a CNN and a vision transformer to identify apple diseases. In their model, the CNN is used to extract the global features of the input image while the transformer extracts local features of the disease region to help the CNN see better. How does it help the CNN? In the case of apple images with different backgrounds (different weather conditions, brightness changes, other objects in the field of view), the CNN could be "distracted" hence the transformer is able to focus the attention of the CNN on the region of interest (the disease region) and not on irrelevant features such as the weather. The use of transformers for automatic zooplankton classification has just begun [@Kyathanahally2022-ma] but it should certainly be explored further in the near future.

@@

```{=latex}
\newpage
```
### The straightforward approach

@@

```{r, uvp traps th234 map}
#| label: fig-uvp-trap-mapping
#| fig-cap: Spatial distribution of (A) sediment traps, (B) $^{234}Th$, (C) UVP5 and (D) their spatio-temporal match-ups.

@@

The idea was based on the assumption that carbon fluxes could be similar in distinct regions of the ocean if those regions had similar environmental conditions thus removing spatial (horizontal) and temporal constraints on match-ups. For this, we tried different combinations of environmental variables such as surface temperature, salinity, nutrients, Chla, $b_{bp}$ or water column properties such as integrated Chla content, oxygen minimum and MLD from monthly climatologies. Then, we used those environmental variables to build a 3D environmental space using a 3D principal component analysis (PCA) decomposition and we projected UVP5 data and trap/$^{234}Th$ in that space based on the environmental characteristics from the respective coordinates and month of each UVP5 and trap/$^{234}Th$ profiles.

@@

This method increased the number of match-ups to \~7800 however the minimization did not find a non ambiguous minimum in the solution space (@fig-solution-space). Instead of having a global minimum, @fig-solution-space shows what looks like a 2D solution plane. In addition, I think the choice of $A$ and $D$ (or $B$, $B = 2D - 1$ ) is not constrained enough. In Figure 5 of @Guidi2008-vp, the light gray color area (low residual error) encompasses a wide region of $A$ and $B$ values where their jackknife procedure found local minima with ($A$, $B$) pairs such as (50, 2) or (150, 5). While $A$ is seen as a constant with no physical meaning, the value of $B$ should really be questioned since it is related to the fractal dimension $D$ of particles and aggregates. Fractal dimensions should not exceed 3 (physical limit) but they have been estimated down to 1.1-1.3 [@Alldredge1988-lx]. In addition, it is possible to obtain coherent POC fluxes by tuning the $A$ value for a fixed $B$ therefore I am not sure that $A$ and $B$ are truly independent.

@@

So far, we have seen that the match-up quest was not enough in the geographical space and not working in the environmental space but what about a method that does not need any match-ups? With XGBoost, it is possible to create a regression model to predict carbon fluxes using environmental variables and geographical variables (depth) as predictors (i.e. explanatory variables) and trap/$^{234}Th$ POC fluxes as target variables. Once the model is trained, it can be applied to the entire data set of PSDs (\> 1.5 million) to derive POC fluxes. We tried to build such a model with a few predictors (i.e. depth, surface temperature, Chla, $b_{bp}$ and the oxygen minimum in the water column). Our model had a $R^{2}$ of 0.47 with a percentage of explained variance dominated by depth (54%), surface Chla (16%), surface temperature (13%) and the oxygen minimum (10%). The minimization of @Guidi2008-vp applied on all true PSDs and predicted POC fluxes led to the same kind of solution space as shown in @fig-solution-space.

Instead of directly predicting carbon fluxes using environmental variable, we could focus on predicting PSD slope or PSD. In a recent paper, @Clements2022-oq used physical and biological variables (see their Table 1) to predict global maps of PSD slope and biovolume at the base of the EZ using UVP5 data and a random forest algorithm. @Clements2022-oq highlighted the importance of Chla (i.e. proxy of biomass), oxygen (i.e. proxy of respiration) and MLD as predictors in their random forest model. They also emphasized the importance of including phytoplankton functional groups and abundance as key controlling factors on both the PSD slope and the biovolume in future models due to their impact on the downward POC flux (i.e. aggregation/disaggregation processes). The study of @Denvil-Sommer2022-oi used both random forest and XGBoost models to predict small and large POC concentrations using geographical variables (depth, latitude and longitude), physical variables (incident light, MLD and temperature), chemical (phosphate and nitrate) and biological variables (Chla and 12 Plankton Functional Types (PFTs) such as diatoms or microzooplankton). @Denvil-Sommer2022-oi showed that for small POC, the most important predictors are the depth level, temperature, microzooplankton and phosphate while the large POC is mostly driven by depth, temperature, MLD, microzooplankton, phaeocystis (another PFT), phosphate and Chla averaged over the MLD. Overall, the results of these studies suggest that it is possible to link the surface environmental and ecosystem structure to the POC distribution in the ocean interior.

@@

I used three methods to derive POC fluxes from UVP5 PSDs. The main issues I encountered were related to the number and accuracy of match-ups as well as the choice of environmental variables to link the surface ocean to the POC distribution below. With the deployment of new BGC-Argo floats equipped with both the UVP6 and the OST as well as the core biogeochemical variables (e.g. Chla, $b_{bp}$, incident light) of the BGC-Argo program, we will hopefully be able to better constraint the PSD-POC relationship however it will probably need a few more years to acquire enough profiles in different oceanic regions to better understand the key drivers of the downward POC flux.

```{=latex}
\newpage
```
## Continuous vertical sequestration

Using the CONVERSE approach, we have estimated that the global carbon sequestration (\> 100 years) flux by the BCP is 0.9-2.6 Pg C y$^{-1}$. This is 2-3 and 3-6 times larger than POC fluxes at, respectively, 1000 and 2000 m. CONVERSE sequestration flux estimates are also up to 4 times larger than those from studies assuming that sequestration occurs below a fixed reference depth [@Honjo2008-kn; @Guidi2015-sj; @Henson2012-ck]. Here below, I highlight some key points and ways to improve our estimations.

@@

Our study on the sequestration flux by the BCP also shows that the contribution of the vertical migration and mixing pumps represent at least 21-34% of the sequestration flux by the BCP. However, those two pumps could probably be better estimated with a multi-instruments approach [@Claustre2021-tx].

@@

With CONVERSE, we estimate that the mixing pump contributes to 11-23% of the sequestration flux by the BCP. However, we were not able to distinct the contribution of the three identified pathways composing the mixing pump due to lacking information in the literature. This could be improved by using the observational framework developed in @Claustre2021-tx by combining satellite, ship-based, mooring and AUV data. BGC-Argo floats have shown that they are adapted to study the detrainment of mixed-layer particles [@DallOlmo2014-ot; @DallOlmo2016-ex] and so have gliders for the eddy subduction pump [@Stukel2017-lc; @Omand2015-mf]. Nonetheless, the study of the large-scale subduction pump will be more challenging because it will require a larger number of measurements made at the scale of ocean basins and over annual cycles [@Claustre2021-tx], hence the need for a multi-instruments approach to better understand it.

@@

With CONVERSE, we estimate that the vertical migration pump contributes to 9-12% (diel) and 0.1-0.4% (seasonal). The estimation of the latter was limited to the northern North Atlantic thus it is clearly underestimated. However, the estimation of vertically migrating (either daily or seasonally) zooplankton and other animals (e.g. fish, jellyfish) is expected to improve with robotic measurements at high temporal resolution and over the entire annual cycle [@Claustre2021-tx]. The combination of acoustic sensors [@Reiss2021-rk; @De_Robertis2021-nw], optical sensors [@Picheral2021-sx] and vertical profiles of biogeochemical variables (e.g. Chla, $b_{bp}$, CDOM) have already shown promising results to detect the migration of mesopelagic organisms in the water column [@Haentjens2020-dc].

```{=latex}
\newpage
```
## Contributions of this work

The embedded classification for AUVs contributes to a better comprehension and quantification of the vertical migration pump. So far, BGC-Argo floats equipped with the UVP6 and the classification model have been deployed in some oceanic regions (e.g. Labrador Sea, Kerguelen Plateau) and have just started to acquire data. In addition to the particle (\> 0.6 mm) identification, the UVP6 also sends their mean area and gray levels which will hopefully contribute to better characterize their morphological traits and therefore better predict their sinking rates (cfr. size versus excess density). The latter are important to estimate the downward carbon flux and the sequestration time.

The BCP can no longer be seen as the sole gravitational settling of particles. Physically and biologically mediated particle injection pumps need to be taken into consideration which is highlighted by our work that estimates their contribution to 21-34% of the total BCP sequestration flux. Our work also shows that the common assumption of sequestration below a fixed reference depth (typically 1000 or 2000 m) misses the bulk of sequestration flux. Therefore, the evaluation of mCDR efficiencies using biologically-driven carbon sequestration fluxes should be estimated over the entire water column.

This work has shown the first comparison of BGC-Argo floats equipped with both the UVP6 and an OST to better quantify carbon fluxes from particle counts and beam attenuation. If the prediction of carbon fluxes at global scale was not a success, this work has highlighted current difficulties and ways to more forward using machine learning approaches. The deployment of BGC-Argo floats in the frame of REFINE with the combination of UVP6 measurements (particle counts and identification), OST beam attenuation fluxes as well as bio-optical variables (e.g. Chla, $b_{bp}$, incident light) will likely improve our understanding of processes that affect the attenuation of the downward carbon flux at local scale.

### 2023-03-24 index.qmd
update

```{=html}
<!--  As of 2022, it has now become obvious to the general public that the Earth's climate is changing. Massive flood events, repeated heatwaves, never-ending droughts, disappearing glaciers, biodiversity collapse, … Reports on weather anomalies were already published in the 1960's, the Intergovernmental Panel on Climate Change (IPCC) was created in 1988, the Kyoto Protocol was adopted in 1997, An Inconvenient Truth was released in 2006, the Paris Agreement was signed in 2015 and still, what has really changed?

As I am writing this part, the COP27 (sponsored by Coca-Cola, no joke) is taking place with an acknowledged record number of fossil fuels lobbyists, bloody stadiums with outside air conditioning are about to be used in the desert for a month and it seems like 'leaders' of the Free World have been having hearing issues. See for yourself:

@@

*"Of course, I would love building new pipelines in your garden"* said the Free World

While I am not an activist (yet?) or a politician (I am too honest to be elected), rest assured, there are no fake news in this work.

I hope you will enjoy it and no, I won't tweet about it. -->
```

>>>>

As of 2022, it has now become obvious to the general public that the Earth's climate is changing. Massive flood events, repeated heatwaves, never-ending droughts, disappearing glaciers, biodiversity collapse, … Reports on weather anomalies were already published in the 1960's, the Intergovernmental Panel on Climate Change (IPCC) was created in 1988, the Kyoto Protocol was adopted in 1997, An Inconvenient Truth was released in 2006, the Paris Agreement was signed in 2015 and still, what has really changed?

As I am writing this part, the COP27 (sponsored by Coca-Cola, no joke) is taking place with an acknowledged record number of fossil fuel lobbyists, bloody stadiums with outside air conditioning are about to be used in the desert for a month and it seems like 'leaders' of the Free World have been having hearing issues. See for yourself:

@@

*"Aaaah okay. Yes you're absolutely right, let's save the banks one more time"* said the Free World

Although I am not an activist (yet?) or a politician (I am too honest to be elected), rest assured that there is no fake news in this work. I hope you will enjoy it.

### 2023-03-24 intro.qmd
update

The most well known pump is the gravitational carbon pump (BGP). It consists in the gravitational settling (i.e. sinking) of POC[^intro-3] to the deep ocean and it is the main contributor (60-70%) of the export flux [@Boyd2019-ez; @Nowicki2022-no]. Historically, the downward POC flux was directly measured using sediment traps [@Martin1987-um; @Honjo2008-kn] and indirectly measured using thorium-234 ($^{234}Th$) deficits as a POC flux proxy [e.g. @Buesseler2006-og]. More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size distributions (PSD, also referred to in the literature as particle size spectra (PSS)) that could then be converted to carbon fluxes [e.g. @Guidi2008-vp].

@@

However, there are still many uncertainties regarding the magnitude of VMP [@Archibald2019-dj; @Aumont2018-wx; @Gorgues2019-qp] and its sequestration potential [@Boyd2019-ez; @Pinti2022-dc] thus its importance for a better understanding of the BCP.

@@

The impact on the downward POC flux of zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2023-qj]. On the one hand, aggregate feeding, filter-feeding, direct ingestion of fecal pellets (i.e. coprophagy) and fragmentation of large aggregates to small slower-sinking particles will likely decrease the downward POC flux. On the other hand, the ingestion of slow-sinking cells to defecation of large and compact (i.e. fast-settling) fecal pellets will increase it. The remaining organic matter (\~15% according to @Turner2015-af) will then be degraded by microbial communities. In the end, only a small fraction of the exported carbon will reach the sediment where it will be stored for millions of years.

@@

The size of aggregates was usually thought to be a key predictor of settling velocity [e.g. @Passow2006-hw] which is coherent with @eq-sinking-speed. However, @Iversen2020-kb did not find evidence for a size-sinking relationship therefore suggesting that excess density would be the main controlling factor. Because the sinking speed determines the time available for degradation in the water column, it also directly impacts the potential for carbon sequestration. With all things considered, there is a thus need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities [@Iversen2020-kb; @Giering2020-zk; @Iversen2023-qj; @Williams2022-xx].

@@

| Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
|-----------------|------------------|-----------------|--------------------|
| Gravitational          | export depth - sediment          | None                      | N/A (142)                            |
| Mixed layer            | 200 - 1,000 m                    | 200 - 1,000 m             | 25 - 100 (54$^{b}$)                  |
| Eddy subduction        | 150 - 400 m                      | 150 - 400 m               | up to 150 (N/A)                      |
| Large-scale subduction | 200 - 1,000 m                    | 200 - 1,000 m             | 25 - 100 (54$^{b}$)                  |
| Diel migration         | 200 - 600 m                      | None                      | up to 250 (150$^{c}$)                |
| Seasonal migration     | 600 - 1,400 m$^{a}$              | None                      | up to 500 (N/A)                      |

: Range of depths generally reached by POC and DOC via the six BCPs [@Claustre2021-tx] and sequestration time potential evaluated by the model of @Boyd2019-ez with an average estimation from the model of @Nowicki2022-no in brackets. None means that a pump does not transport DOC. N/A means that the pump was not evaluated in the model. (a) 1,400 m corresponds to the overwintering depth of *Calanus finmarchicus* as observed by @Jonasdottir2015-ob. (b) 54 years depicts the mean sequestration time of the mixing pump, without the contribution of the eddy subduction pump (c) Migration of mesopelagic fish was neglected in the model. {#tbl-PumpDepthRange}

>>>>

The most well known pump is the gravitational carbon pump (BGP). It consists in the gravitational settling (i.e. sinking) of POC[^intro-3] to the deep ocean and it is the main contributor (60-70%) of the export flux [@Boyd2019-ez; @Nowicki2022-no]. Historically, the downward POC flux was directly measured using sediment traps [@Martin1987-um; @Honjo2008-kn] and indirectly measured using thorium-234 ($^{234}Th$) deficits as a POC flux proxy [e.g. @Buesseler2006-og]. More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size distributions (PSDs, also referred to in the literature as particle size spectra (PSS)) that could then be converted to carbon fluxes [e.g. @Guidi2008-vp].

@@

However, there are still many uncertainties regarding the magnitude of the VMP [@Archibald2019-dj; @Aumont2018-wx; @Gorgues2019-qp] and its sequestration potential [@Boyd2019-ez; @Pinti2023-td] thus its importance for a better understanding of the BCP.

@@

The impact on the downward POC flux by zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2023-qj]. On the one hand, aggregate feeding, filter-feeding, direct ingestion of fecal pellets (i.e. coprophagy) and fragmentation of large aggregates to small slower-sinking particles will likely decrease the downward POC flux. On the other hand, the ingestion of slow-sinking cells to defecation of large and compact (i.e. fast-settling) fecal pellets will increase it. The remaining organic matter (\~15% according to @Turner2015-af) will then be degraded by microbial communities. In the end, only a small fraction of the exported carbon will reach the sediment where it will be stored for millions of years.

@@

The size of aggregates was usually thought to be a key predictor of settling velocity [e.g. @Passow2006-hw] which is coherent with @eq-sinking-speed. However, @Iversen2020-kb did not find evidence for a size-sinking relationship therefore suggesting that excess density would be the main controlling factor. Because the sinking speed determines the time available for degradation in the water column, it also directly impacts the potential for carbon sequestration. With all things considered, there is thus a need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities [@Iversen2020-kb; @Giering2020-zk; @Iversen2023-qj; @Williams2022-xx].

@@

| Pump                   | POC export/injection depth range | DOC injection depth range | Sequestration time potential (years) |
|-----------------|-------------------|-----------------|---------------------|
| Gravitational          | export depth - sediment          | None                      | N/A (142)                            |
| Mixed layer            | 200 - 1,000 m                    | 200 - 1,000 m             | 25 - 100 (54$^{b}$)                  |
| Eddy subduction        | 150 - 400 m                      | 150 - 400 m               | up to 150 (N/A)                      |
| Large-scale subduction | 200 - 1,000 m                    | 200 - 1,000 m             | 25 - 100 (54$^{b}$)                  |
| Diel migration         | 200 - 600 m                      | None                      | up to 250 (150$^{c}$)                |
| Seasonal migration     | 600 - 1,400 m$^{a}$              | None                      | up to 500 (N/A)                      |

: Range of depths generally reached by POC and DOC via the six BCPs [@Claustre2021-tx] and sequestration time potentials evaluated by the model of @Boyd2019-ez with an average estimation from the model of @Nowicki2022-no in brackets. None means that a pump does not transport DOC. N/A means that the pump was not evaluated in the model. (a) 1,400 m corresponds to the overwintering depth of *Calanus finmarchicus* as observed by @Jonasdottir2015-ob. (b) 54 years depicts the mean sequestration time of the mixing pump, without the contribution of the eddy subduction pump (c) Migration of mesopelagic fish was neglected in the model. {#tbl-PumpDepthRange}

### 2023-03-24 merci.qmd
update

remerciement

thomas boniface pour les figures

anais

alexbarth pour son bouquin sur le machine car j'ai passé mes 3 première semaines à faire ça vu que j'y connaissais absolument rien

faire des stats du vélo (et meme mettre une carte si possible rapidement)

tous ceux qui ont partagé des moments sportifs ensembles, au bloc mais surtout en vélo et qui malgré la souffrance acceptaient encore de venir rouler avec moi (Ophélie, Flavien, Thelma, Laetitia, Alexandre, Satoshi)

Julien, des challenges à n'en plus finir, un peu mazo au final, à force d'aller en enfer, tu incarneras le diable en personne bientôt(#Ducasse)

Les montois (Mathieu, inconditionel, Florent, les Juliens, Chris)

Les co-Phd : Chloé, AMP (the best), Anais

Hadrien who accepted to be my roomate in Liège

alice

marine of course

expatrié que je suis et qui venu me voir 'ou profiter, I'll never really know)

les Liégeois (Thomas, Nico, Hadri, Flop, Florent)

Lucie, mon old roomate (lucie lelotte)

Alex, my 15th french history roomate, I am glad you were there suring the 6 pm curfew we had and

Les montois bien sûr

tous xeux qui sont descendus me voir, (vous veniez pour moi ou la mer tirquoise, j'ai des doutes mtn)

>>>>

# Acknowledgments

Thanks to the members of the jury for taking the time to read and evaluate this three and a half year work. I am looking forward to debate it during the defense.

Cette thèse a été particulière à bien des égards, avec son lot de positifs et de négatifs mais surtout avec son lot de rebondissements qui m'ont permis d'affûter mes connaissances, ma résilience et je l'espère ma tolérance. Pour cela, je tiens à remercier chaleureusement toutes les personnes y ayant contribué de près ou de loin.

Merci à Hervé, Lionel et Jean-Olivier pour leur accueil au Laboratoire de Villefranche-sur-Mer et pour leur disponibilité ainsi que pour leurs conseils avisés, tant scientifiques que personnels.

Merci à Alex, Charles et Arthur pour leur aide et leur gentillesse.

Merci également aux personnes que j'ai pu cotoyer à COMPLEX comme chez OMTAB depuis mon passage en stage en 2018 et les multiples allers-retours Belgique-France.

Merci à Antoine Mangin de m'avoir prêté un bureau dans les locaux d'ACRI sur Grasse l'été dernier pendant que mon appartement servait d'Airbnb estival pour le plus grand bonheur du portefeuille de mon proprio.

Un grand merci à mes colocs de bureau: Ophélie, Laetitia et Thelma pour votre soutien et votre bonne humeur sans faille (contrairement à moi).

Merci à AMP (elle se reconnaitra), Chloé et Anaïs pour les pauses midi sur le ponton. AMP, merci d'avoir accepté de vivre virtuellement avec moi afin que la sécu daigne enfin me donner une carte vitale !

Ne me demandez pas des noms de restos ou de bars sur la côte d'Azur, je n'en ai aucune idée. Par contre, je connais pas mal les routes du coin. Merci à tous ceux qui ont partagé (ou subi) des kilomètres avec moi, surtout en montée (le plat pays, qui est le mien): Flavien, Ophé, Lucas, Thelma, Laeti et Satoshi.

Un grand merci à Marine et Alice pour leur bienveillance. Marine, je te souhaite le meilleur pour tes projets futurs mais fais attention à qui tu donnes tes clés quand même.

Merci également à Alex qui a été un excellent roommate pendant les confinements, couvre-feux et ordonnances en tout genre. See you in London !

Merci aux Liégeois(es) (Thomas, Florent, Flop, Nico, Evgeny et Lucie) et en particulier Hadrien qui a accepté de m'avoir comme coloc pendant plusieurs mois (et qui y a survécu).

Merci aux Montois (Mathieu, Chris, Flo, Ju & Ju) pour votre soutien à distance, un mariage en 3 actes et pour vos visites sur Nice. Petite dédicace à Ju K qui, à force d'avoir joué avec le diable en subissant l'enfer des pourcentages, va littéralement en devenir un. Vivement 2024 pour casser la baraque.

Un très grand merci à ma famille nucléaire qui ne comprend toujours pas mon job mais qui fait semblant quand je lui explique, c'est le principal. C'est vrai que la défense tombe mal au niveau timing donc on se voit à la Ducasse :-)

Enfin, un grand merci à une personne toute particulière qui en a probablement marre que je parle de la Belgique mais qui à présent est parfaite bilingue, mission accomplie?

### 2023-03-24 observingCarbonCycle.qmd
update

The UVP6-LP (hereafter referred to as UVP6) is thus both a particle counter (0.1 mm - 2 mm) and a particle classifier (\> 0.6 mm). The former counts particles in different class sizes in order to compute the PSD while the latter extracts pixel-related and geometric features from each image to classify them (i.e. assign them a taxonomic group) directly on board using artificial intelligence (@sec-uvp-ia) with the aim of estimating zooplankton biomass [@Claustre2020-jo; @Claustre2021-tx].

@@

1.  BGC-Argo floats equipped with a UVP6 and the zooplankton recognition algorithm are able to classify particles larger than 0.6 mm. However, these AUVs are limited in terms of their battery life. They are typically deployed for several years, which requires special attention to the energy consumption of each sensor. In @sec-uvp6, I will therefore describe the methodology used to develop the zooplankton classification model for the UVP6 accounting for energy limitations.
2.  Four BGC-Argo floats equipped with both an OST and the UVP6 were deployed for the first time in the Labrador Sea in late May 2022 to study the dynamics of carbon fluxes and particle size distributions. In this chapter, I will explore the seasonality of the later and compare how the two sensors relate as two independent measurements of particle and carbon fluxes.
3.  Many studies have assumed that carbon sequestration only occurs in the deep ocean (typically 1000 m). In @bcp-sequestration, we reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.

>>>>

The UVP6-LP (hereafter referred to as UVP6) is thus both a particle counter (0.1 mm - 2 mm) and a particle classifier (\> 0.6 mm). The former counts particles in different class sizes in order to compute the PSD while the latter extracts pixel-related and geometric features from each image to classify them (i.e. assign them a taxonomic group) directly on board using artificial intelligence (@sec-uvp6) with the aim of estimating zooplankton biomass [@Claustre2020-jo; @Claustre2021-tx].

@@

1.  BGC-Argo floats equipped with a UVP6 and the zooplankton recognition algorithm are able to classify particles larger than 0.6 mm. However, these AUVs are limited in terms of their battery life. They are typically deployed for several years, which requires special attention to the energy consumption of each sensor. In @sec-uvp6, I will therefore describe the methodology used to develop the zooplankton classification model for the UVP6 accounting for energy limitations.
2.  Four BGC-Argo floats equipped with both an OST and the UVP6 were deployed for the first time in the Labrador Sea in late May 2022 to study the dynamics of carbon fluxes and particle size distributions. In this chapter, I will explore the seasonality of the later and compare how the two sensors relate as two independent measurements of particle and carbon fluxes.
3.  Many studies have assumed that carbon sequestration only occurs in the deep ocean (typically 1000 m). In @sec-bcp-sequestration, we reassess the global sequestration ($\geq$ 100 years) of the BCP and its components by using a continuous approach to carbon storage as opposed to the common view of long-term sequestration below a fixed reference depth.
